On Complexity Issue of Online Learning Algorithms

### 2020-08-30 publications/complexity-it[04.04].tex
Add files via upload

\author{Yuan~Yao
\thanks{Manuscript submitted March 28, 2005, with Confirmation Number 1630. This work was supported by NSF grant 0325113.}
\thanks{Yuan Yao is with the Department of Mathematics, University of California at Berkeley, Berkeley, CA 94720 (E-mail: yao@math.berkeley.edu). This work
was done while the author visited the Toyota Technological Institute at Chicago, 1427 East 60th Street, Chicago, IL 60637.}}

@@

It is possible to improve the rate in $\lambda$ by turning back to Markov's Inequality. In fact, the reason of loss in $\lambda$ lies in the application
of the Hoeffding-style inequalities which, compared to Markov's Inequality, replace the variance by its uniform upper bounds.
Based on this observation, we obtain the following result (as Theorem B*) for the averaging process by using Markov's Inequality,
\[ \|\bar{f}_t - f^\ast_\lambda\|_K \leq O(\lambda^{-1}t^{-1/2} \delta^{-1/2}), \ \ \mbox{$\theta \in [0,1)$}, \]
which holds with probability at least $1-\delta$ ($\delta\in (0,1)$). However, an explicit upper bound on variance only leads to
$O(\lambda^{-3/2}t^{-1/2} \delta^{-1/2})$. It is not clear yet whether we can improve this by using other variance-based inequalities, such as
Bennet's or Bernstein's Inequality.

@@

\begin{mainthm}[Theorem B]
Let $\lambda\leq \lambda_0$, $\gamma_t = t^{-\theta}/(C^2_K + \lambda)$ for some $\theta\in [0,1)$, and $f_1=0$.
Then for all $t\in \N$ there holds for (\ref{eq:averaging})
\[ \| \bar{f}_t - f^\ast_\lambda \|_K \leq \Err_{init}(t) + \Err_{samp}(t). \]
where
\[ \Err_{init}(t)\leq  C_1 \left( \frac{1}{\lambda t} \right), \]
and with probability at least $1-\delta$ ($\delta\in (0,1)$),
\[ \Err_{samp}(t) \leq C_2 \left(\frac{1}{\lambda} \right)^2 \sqrt{\frac{1}{t}} \log^{1/2}\frac{2}{\delta}. \]
Here $C_1=D_\theta(\lambda_0+C^2_K) \|f^\ast_\lambda\|_K$ and $C_2=8 D_\theta C_K M_\rho(\lambda_0+C^2_K)$.
\end{mainthm}

@@

Let $\sigma^2_\lambda=\E[\|(y - f^\ast_\lambda(x)) - \lambda f^\ast_\lambda(x)\|^2$ for some $\sigma_\lambda \geq 0$. Then by Markov's Inequality we
improve the rate in $\lambda$ as in the following theorem, whose proof will be given in Section VI.

\begin{mainthm}[Theorem B*]
Let $\lambda\leq \lambda_0$, $\gamma_t = t^{-\theta}/(C^2_K + \lambda)$ for some $\theta\in [0,1)$, and $f_1=0$.
Then the following holds for all $t\in \N$,
\[ \| \bar{f}_t - f^\ast_\lambda \|_K \leq \Err_{init}(t) + \Err_{samp}(t), \]
where
\[ \Err_{init}(t)\leq D_\theta(\lambda_0+C^2_K) \|f^\ast_\lambda\|_K \left( \frac{1}{\lambda t} \right), \]
and with probability at least $1-\delta$ ($\delta\in (0,1)$),
\[ \Err_{samp}(t) \leq \frac{D_\theta \sigma_\lambda}{ \lambda \sqrt{\delta t}}. \]
\end{mainthm}

\begin{rem} It should be noticed that $\sigma_\lambda$ depends on $\lambda$, whence it is not clear yet if the rate $O(\lambda^{-1}t^{-1/2})$ is
achievable. A preliminary result by Proposition \ref{prop:varbnd}-2, gives an estimate $\sigma_\lambda=O(\lambda^{-1/2})$.
With this we have
\[ \|\bar{f}_t - f^\ast_\lambda\|_K \leq O\left(\left(\frac{1}{\lambda}\right)^{3/2}\sqrt{\frac{1}{\delta  t}}\right). \]
\end{rem}

@@

\begin{rem}
A recent result \cite{CapDev05} shows that the rate $O(\lambda^{-1} t^{-1/2})$ is optimal in the sense that it leads to a
convergence rate asymptotically meeting the minimax lower bound.
Theorem B tells us that the averaging process achieves $O(\lambda^{-2} t^{-1/2})$,
which is optimal in $t$ but suboptimal in $\lambda$. Theorem B* improves
this to $O(\lambda^{-1} t^{-1/2})$, for uniformly bounded $\sigma^2_\lambda$; while for the estimate
provided in Proposition \ref{prop:varbnd}-2, one only achieves $O(\lambda^{-3/2} t^{-1/2})$. Yet it is an open
problem if we can achieve the optimal rate in both $\lambda$ and $t$.
\end{rem}

@@

\begin{thm} \label{thm:average}
Let $\DS \gamma_t = t^{-\theta}/\amax$ ($\theta\in [0,1)$) and $w_1=0$. Then for all $t\in \N$ there holds
\[ \| \bar{w}_t -w^\ast \| \leq \Err_{init}(t) + \Err_{samp}(t), \]
where
\[ \Err_{init}(t)\leq  D_\theta \left(\frac{1}{\alpha t} \right)\|r_1\| , \]
and with probability at least $1-\delta$,
\[ \Err_{samp}(t) \leq  \frac{8 D_\theta \beta}{\amax} \left(\frac{1}{\alpha}\right)^2 \sqrt{\frac{1}{t}} \log^{1/2} \frac{2}{\delta}. \]
\end{thm}

@@

\begin{prop} For all $t\in \N$,
\begin{equation} \label{eq:total}
r_{t} = \hat{\Pi}_1^{t-1} r_1 -
\sum_{k=1}^{t-1}\gamma_k  \Hat{\Pi}_{k+1}^{t-1} \chi_k,
\end{equation}
where $\chi_k=(A_k - \hat{A}) w_k + \hat{B}-B_t$ ($1\leq k \leq t$).
\end{prop}

@@

2. $\left\|\sum_{j=k}^t \Pi_{k+1}^j \right\| \leq D_\theta \alpha^{-1} k^\theta$, and the same holds for $\|\sum_{j=k}^t\hat{\Pi}_{k+1}^t\|$;

@@

2. Note that
\[ \left\|\sum_{j=k}^t \Pi_{k+1}^j \right\| \leq \sum_{j=k}^t \prod_{i=k+1}^j \left( 1- \frac{\alpha}{i} \right), \]
whence the result follows from Lemma \ref{lem:estimates}-2.

@@

By Proposition \ref{prop:bounds}-2 with $k=1$,
\[ \left\|\sum_{j=1}^{t-1} \hat{\Pi}_1^j \right\| \leq \frac{D_\theta}{\alpha}, \]
which gives the first bound.

As to the second bound, by Proposition \ref{prop:bounds}-2 and Proposition \ref{prop:bounds}-6 with $w_1=0$, we obtain
\begin{eqnarray}
\|\eta_k\| & \leq & \frac{\gamma_k}{t} \left\| \sum_{j=k}^{t} \hat{\Pi}_{k+1}^j \right\| \|\chi_k\| \nonumber \\
& \leq & \frac{8 \beta D_\theta \amax}{t \amin^2} = c_\eta. \label{eq:eta}
\end{eqnarray}
Applying Pinelis-Hoeffding inequality (Lemma \ref{lem:pinelis}),
\[ \Prob \left\{ \left\|\sum_{k=1}^{t-1} \eta_k \right\| \geq \epsilon \right\} \leq
2 \exp \left\{-\frac{\epsilon^2}{2\sum_{k=1}^{t-1} c_\eta^2} \right\},  \]
and setting the right hand to be $\delta$, we obtain
\[ \epsilon \leq  \sqrt{2t} c_\eta \log^{1/2} \frac{2}{\delta}.\]
The second bound follows from (\ref{eq:eta}).
\end{pf}

@@

\begin{lem} For all $t\in \N$,
\[ \bar{r}_{t} = \frac{1}{t} \left(\sum_{j=1}^{t-1} \Pi_1^j \right) r_1 - \sum_{k=1}^{t-1}\frac{\gamma_k}{t}  \left(\sum_{j=k}^{t-1}\Pi_{k+1}^{j}\right) Y_k. \]
\end{lem}

@@

\begin{thm} \label{thm:B1}
Let $\gamma_t = t^{-\theta}/\amax$ ($\theta\in [0,1)$) and $w_1=0$. Define $\alpha=\amin/\amax \in (0,1]$.
Then the following holds for all $t\in \N$,
\[ \| \bar{w}_t -w^\ast \| \leq \Err_{init}(t) + \Err_{samp}(t). \]
Here
\[ \Err_{init}(t)\leq  D_\theta\left(\frac{1}{\alpha t} \right)\|r_1\| , \]
and with probability at least $1-\delta$ ($\delta\in (0,1)$),
\[ \Err_{samp}(t) \leq  \frac{D_\theta \sigma}{\sqrt{\delta} \amax} \left(\frac{1}{\alpha}\right) \sqrt{\frac{1}{t}}. \]
\end{thm}

\begin{pf}
The initial error bound follows from Proposition \ref{prop:bounds}-2. As to the sample error, note that
\begin{eqnarray*}
\E \|\sum_{k=1}^{t-1} \eta_k\|^2 & \leq & \sum_{k=1}^{t-1} \frac{\gamma^2_k}{t^2}  (\psi^0_\theta(t,k,\alpha))^2 \E \|Y_k\|^2 \\
& \leq & \frac{D_\theta^2 \sigma^2}{\amax^2\alpha^2}t^{-1}.
\end{eqnarray*}
where in the second last step,
\[ \psi^0_\theta(t,k,\alpha) = \sum_{j=k}^{t-1} \prod_{i=k+1}^j \left(1-\frac{\alpha}{i^\theta} \right), \]
and the last is due to Lemma \ref{lem:estimates}-2 and Finiteness Condition-C. The sample error bound then follows from Markov inequality
by regarding $X=\|\sum_{k=1}^{t-1} \eta_k\|^2$.
\end{pf}

@@

The following proposition gives an estimate of $\sigma_\lambda$ via $\|f^\ast_\lambda\|_K$.

@@

2. $\sigma_\lambda \leq M_\rho(1+\sqrt{\lambda}+\lambda)/\sqrt{\lambda}$.
\end{prop}

\begin{pf}
1. Note that
\[ f^\ast_\lambda = \arg \min_{f\in \H_K} \|f - f_\rho\|_{\L^2_\rho}^2 + \lambda \|f\|^2_K. \]
Taking $f=0$, we have
\[ \|f^\ast_\lambda - f_\rho\|_{\L^2_\rho}^2 + \lambda \|f^\ast_\lambda\|^2_K \leq \|f_\rho\|_{\L^2_\rho}^2 \leq M_\rho^2, \]
which leads to the result.

2. By definition
\begin{eqnarray*}
\sigma_\lambda & = & \sqrt{\E\|(y-f^\ast_\lambda(x))-\lambda f^\ast_\lambda\|_K^2} \\
& \leq & \E \|y-f^\ast_\lambda(x)\|_K + \lambda \|f^\ast_\lambda\|_K \\
& \leq &  M_\rho+ \frac{M_\rho}{\sqrt{\lambda}} + M_\rho \sqrt{\lambda}
\end{eqnarray*}
which ends the proof.
\end{pf}

@@

In this paper, we have shown by probabilistic upper bounds that a two-stage online learning algorithm,
the stochastic approximation of the gradient descent method followed by an averaging process, can achieve
the almost sure convergence with an optimal asymptotic rate with respect to the sample size, as good as ``batch learning''.
Moreover considering the regularization parameter and confidence, the best results obtained so far are,
$O(\lambda^{-2} t^{-1/2} \log^{1/2} 1/\delta)$ (Theorem B) or $O(\lambda^{-3/2} t^{-1/2} \delta^{-1/2})$ (Remark 2.4 for Theorem B*)
under the conditions in this paper, and $O(\lambda^{-1} t^{-1/2} \delta^{-1/2})$ (Theorem B*) under a strong condition on variance.

@@

2. For $\alpha\in (0,1]$, $\theta\in [0,1)$, and all $t\in \N$,
\[ \psi^0_\theta(t,k,\alpha):=\sum_{j=k}^{t-1} \prod_{i=k+1}^j \left(1 - \frac{\alpha}{i^\theta} \right)
\leq \frac{D_\theta}{\alpha} k^{\theta};
\]

@@

2. Notice that
\[ \psi^0_\theta(t,\alpha)=1+ \sum_{j=k+1}^{t-1} \prod_{i=k+1}^j \left(1 - \frac{\alpha}{i^\theta} \right)
\leq \frac{D_\theta}{\alpha} k^{\theta}. \]
As $1\leq k^\theta/\alpha$, it suffices to prove the second term is bounded by $(D_{\theta}-1)k^\theta/\alpha$.

Let $\alpha'=\alpha/(1-\theta)$. Using part 1 with $p=1$, we obtain
\begin{eqnarray*}
\prod_{i=k+1}^j \left(1 - \frac{\alpha}{i^\theta} \right) & \leq &  e^{\alpha'[(k+1)^{1-\theta} - (j+1)^{1-\theta}]},
\end{eqnarray*}
whence
\begin{eqnarray*}
& & \sum_{j=k+1}^{t-1} \prod_{i=k+1}^j \left(1 - \frac{\alpha}{i^\theta} \right) \\
& \leq & \sum_{j=k+1}^{t-1} e^{\alpha'[(k+1)^{1-\theta} - (j+1)^{1-\theta}]} \\
& \leq & e^{\alpha'(k+1)^{1-\theta}}\int_{k+1}^\infty e^{- \alpha' x^{1-\theta}} d x \\
& \leq & (\alpha')^{-1/(1-\theta)} C_\theta' [(\alpha')^{1/(1-\theta)} (k+1)]^{\theta}, \\
& & \ \ \ \ \ \mbox{(by Lemma \ref{lem:key2})} \\
& \leq & 2^{\theta/(1-\theta)} \left\{ 1+\Gamma\left(\frac{1}{1-\theta}\right)\right\} \left(\frac{1}{\alpha}\right)k^\theta \\
& = & \frac{D_\theta-1}{\alpha} k^\theta,
\end{eqnarray*}
as desired.

>>>>

\author{Yuan~Yao
\thanks{Manuscript submitted March 28, 2005, with CLN \# 5-215. Revised April 4, 2005. This work was supported by NSF grant 0325113.}
\thanks{Yuan Yao is with the Department of Mathematics, University of California at Berkeley, Berkeley, CA 94720 (E-mail: yao@math.berkeley.edu). This work
was done while the author visited the Toyota Technological Institute at Chicago, 1427 East 60th Street, Chicago, IL 60637.}}

@@

It is possible to improve the rate in $\lambda$ by turning back to Markov's Inequality. In fact, the reason of loss in $\lambda$ lies in the application
of the Hoeffding-style inequalities which, compared to Markov's Inequality, replace the variance by its uniform upper bounds.
Based on this observation, we obtain the following result (as Theorem B*) for the averaging process by using Markov's Inequality,
\[ \|\bar{f}_t - f^\ast_\lambda\|_K \leq O(\lambda^{-1}t^{-1/2} \delta^{-1/2}), \ \ \mbox{$\theta \in [0,1)$}, \]
which holds with probability at least $1-\delta$ ($\delta\in (0,1)$). We conjecture that this can be improved to be 
$O(\lambda^{-1} t^{-1/2} \log^{1/2}1/\delta)$ by using other variance-based inequalities, such as Bennet's or Bernstein's Inequality.

@@

\begin{mainthm}[Theorem B]
Let $\lambda\leq \lambda_0$, $\gamma_t = t^{-\theta}/(C^2_K + \lambda)$ for some $\theta\in [0,1)$, and $f_1=0$.
Then for all $t\in \N$ there holds for (\ref{eq:averaging})
\[ \| \bar{f}_t - f^\ast_\lambda \|_K \leq \Err_{init}(t) + \Err_{samp}(t). \]
where
\[ \Err_{init}(t)\leq  C_1 \left( \frac{1}{\lambda t} \right), \]
and with probability at least $1-\delta$ ($\delta\in (0,1)$),
\[ \Err_{samp}(t) \leq C_2 \left(\frac{1}{\lambda} \right)^2 \sqrt{\frac{1}{t}} \log^{1/2}\frac{2}{\delta}. \]
Here $C_1=D_\theta(\lambda_0+C^2_K) \|f^\ast_\lambda\|_K$ and $C_2=2^{3+\theta} D_\theta C_K M_\rho(\lambda_0+C^2_K)$.
\end{mainthm}

@@

The rate in $\lambda$ can be improved. 
Let $\sigma^2_\lambda=\E[\|(y - f^\ast_\lambda(x))K_x - \lambda f^\ast_\lambda\|_K^2$ for some $\sigma_\lambda \geq 0$. By Markov's Inequality we
obtain the following theorem, whose proof will be given in Section VI.

\begin{mainthm}[Theorem B*]
Let $\lambda\leq \lambda_0$, $\gamma_t = t^{-\theta}/(C^2_K + \lambda)$ for some $\theta\in [0,1)$, and $f_1=0$.
Then the following holds for all $t\in \N$,
\[ \| \bar{f}_t - f^\ast_\lambda \|_K \leq \Err_{init}(t) + \Err_{samp}(t), \]
where
\[ \Err_{init}(t)\leq D_\theta(\lambda_0+C^2_K) \|f^\ast_\lambda\|_K \left( \frac{1}{\lambda t} \right), \]
and with probability at least $1-\delta$ ($\delta\in (0,1)$),
\[ \Err_{samp}(t) \leq \frac{2^\theta D_\theta \sigma_\lambda}{ \lambda \sqrt{\delta t}}. \]
\end{mainthm}

\begin{rem} Proposition \ref{prop:varbnd}-3 gives an estimate on $\sigma_\lambda$,
\[ \sigma_\lambda \leq M_\rho \sqrt{5(\lambda_0+C^2_K)}. \]
\end{rem}

@@

\begin{rem}
A recent result \cite{CapDev05} shows that the rate $O(\lambda^{-1} t^{-1/2})$ is optimal in the sense that it leads to a
convergence rate asymptotically meeting the minimax lower bound.
Theorem B tells us that the averaging process achieves $O(\lambda^{-2} t^{-1/2})$,
which is optimal in $t$ but suboptimal in $\lambda$. Theorem B* improves
this to $O(\lambda^{-1} t^{-1/2})$, though it only leads to convergence in probability. 
\end{rem}

@@

\begin{thm} \label{thm:average}
Let $\DS \gamma_t = t^{-\theta}/\amax$ ($\theta\in [0,1)$) and $w_1=0$. Then for all $t\in \N$ there holds
\[ \| \bar{w}_t -w^\ast \| \leq \Err_{init}(t) + \Err_{samp}(t), \]
where
\[ \Err_{init}(t)\leq  D_\theta \left(\frac{1}{\alpha t} \right)\|r_1\| , \]
and with probability at least $1-\delta$,
\[ \Err_{samp}(t) \leq  \frac{2^{3+\theta} D_\theta \beta}{\amax} \left(\frac{1}{\alpha}\right)^2 \sqrt{\frac{1}{t}} \log^{1/2} \frac{2}{\delta}. \]
\end{thm}

@@

\begin{prop} For all $t\in \N$,
\begin{equation} \label{eq:total}
r_{t} = \hat{\Pi}_1^{t-1} r_1 -
\sum_{k=1}^{t-1}\gamma_k  \Hat{\Pi}_{k+1}^{t-1} \chi_k,
\end{equation}
where $\chi_k=(A_k - \hat{A}) w_k + \hat{B}-B_k$ ($1\leq k \leq t$).
\end{prop}

@@

2. For all integers $k\in [0,t]$,
\[
\left\|\sum_{j=k}^t \Pi_{k+1}^j \right\| \leq \left\{
\begin{array}{lr}
2^\theta D_\theta \alpha^{-1} k^\theta, & 1\leq k \leq t; \\
D_\theta \alpha^{-1}, & k=0.
\end{array}
\right.
\]
The same also holds for $\|\sum_{j=k}^t\hat{\Pi}_{k+1}^t\|$;

@@

2. By Lemma \ref{lem:estimates}-2,
\[ \left\|\sum_{j=k}^t \Pi_{k+1}^j \right\| \leq 1+\sum_{j=k+1}^t \prod_{i=k+1}^j \left( 1- \frac{\alpha}{i} \right)\leq 1+\frac{D_\theta-1}{\alpha}(k+1)^\theta, \]
where if $k=0$, $r.h.s. \leq D_\theta \alpha^{-1}$, and if $k\geq 1$, $r.h.s. \leq 2^{\theta} D_\theta \alpha^{-1} k^{\theta}$.

@@

The initial error bound follows from Propoistion \ref{prop:bounds}-2 with $k=0$.
As to the sample error bound, by Proposition \ref{prop:bounds}-2 and Proposition \ref{prop:bounds}-7 with $w_1=0$, we obtain
\begin{eqnarray}
\|\eta_k\| & \leq & \frac{\gamma_k}{t} \left\| \sum_{j=k}^{t-1} \hat{\Pi}_{k+1}^j \right\| \|\chi_k\| \nonumber \\
& \leq & \frac{2^{\theta+3} \beta D_\theta \amax}{t \amin^2} = c_\eta. \label{eq:eta}
\end{eqnarray}
Applying Pinelis-Hoeffding inequality (Lemma \ref{lem:pinelis}),
\[ \Prob \left\{ \left\|\sum_{k=1}^{t-1} \eta_k \right\| \geq \epsilon \right\} \leq
2 \exp \left\{-\frac{\epsilon^2}{2\sum_{k=1}^{t-1} c_\eta^2} \right\},  \]
and setting the right hand to be $\delta$, we obtain
\[ \epsilon \leq  \sqrt{2t} c_\eta \log^{1/2} \frac{2}{\delta}.\]
The second bound follows from (\ref{eq:eta}).
\end{pf}

@@

\begin{lem} For all $t\in \N$,
\[ \bar{r}_{t} = \frac{1}{t} \left(\sum_{j=0}^{t-1} \Pi_1^j \right) r_1 - \sum_{k=1}^{t-1}\frac{\gamma_k}{t}  \left(\sum_{j=k}^{t-1}\Pi_{k+1}^{j}\right) Y_k. \]
\end{lem}

@@

\begin{thm} \label{thm:B1}
Let $\gamma_t = t^{-\theta}/\amax$ ($\theta\in [0,1)$) and $w_1=0$. Define $\alpha=\amin/\amax \in (0,1]$.
Then the following holds for all $t\in \N$,
\[ \| \bar{w}_t -w^\ast \| \leq \Err_{init}(t) + \Err_{samp}(t). \]
Here
\[ \Err_{init}(t)\leq  D_\theta\left(\frac{1}{\alpha t} \right)\|r_1\| , \]
and with probability at least $1-\delta$ ($\delta\in (0,1)$),
\[ \Err_{samp}(t) \leq  \frac{2^{\theta} D_\theta \sigma}{\sqrt{\delta} \amax} \left(\frac{1}{\alpha}\right) \sqrt{\frac{1}{t}}. \]
\end{thm}

\begin{pf}
The initial error bound follows from Propoistion \ref{prop:bounds}-2 with $k=0$.

As to the sample error, note that
\begin{eqnarray*}
\E \|\sum_{k=1}^{t-1} \eta_k\|^2 & \leq & \sum_{k=1}^{t-1} \frac{\gamma^2_k}{t^2}  \E\|\sum_{j=k}^{t-1}\Pi_{k=k+1}^j\|^2 \E \|Y_k\|^2 \\
& \leq & \frac{2^{2\theta} D_\theta^2 \sigma^2}{\amax^2\alpha^2}t^{-1}.
\end{eqnarray*}
where the last is due to Proposition \ref{prop:bounds}-2 and Finiteness Condition-C. The sample error bound then follows from Markov inequality
by taking $X=\|\sum_{k=1}^{t-1} \eta_k\|^2$.
\end{pf}

@@

The following proposition gives an estimate of $\sigma_\lambda$.

@@

2. $\|f^\ast_\lambda\|_{\L^2_\rho} \leq 2 M_\rho$;

3. $\sigma_\lambda \leq M_\rho\sqrt{5(\lambda+C^2_K)}$.
\end{prop}

\begin{pf}
1. Note that
\[ f^\ast_\lambda = \arg \min_{f\in \H_K} \|f - f_\rho\|_{\L^2_\rho}^2 + \lambda \|f\|^2_K. \]
Taking $f=0$, we have
\begin{equation} \label{eq:varbnd1}
\|f^\ast_\lambda - f_\rho\|_{\L^2_\rho}^2 + \lambda \|f^\ast_\lambda\|^2_K \leq \|f_\rho\|_{\L^2_\rho}^2 \leq M_\rho^2, 
\end{equation}
which leads to the result.

2. From (\ref{eq:varbnd1}), we obtain $\|f^\ast_\lambda - f_\rho\|_{\L^2_\rho} \leq M_\rho$. The result then follows from
\begin{eqnarray*}
\|f_\lambda^\ast\|_{\L^2_\rho} & \leq & \| f_\lambda^\ast - f_\rho \|_{\L^2_\rho} + \|f_\rho \|_{\L^2_\rho} \leq 2M_\rho. 
\end{eqnarray*}

3. Note that $\E[yK_x]=L_K f_\rho$ and $\E[f^\ast_\lambda(x)K_x]=L_K f^\ast_\lambda$. Then
\begin{eqnarray*}
\sigma_\lambda^2 & = & \E\|(y-f^\ast_\lambda(x))K_x-\lambda f^\ast_\lambda\|_K^2 \\
& = & \E [(f^\ast_\lambda(x)-y)^2 K(x,x)] \\
& & \ \ \ \ \ + 2 \lambda \<f^\ast_\lambda, L_K (f^\ast_\lambda-f_\rho)\>_K + \lambda^2 \|f^\ast_\lambda\|_K^2 
\end{eqnarray*}
where
\begin{eqnarray*}
& & \E[(f^\ast_\lambda(x)-y)^2 K(x,x)] \\
& \leq & C^2_K (\|f^\ast_\lambda-f_\rho\|^2_{\L^2_\rho}+ \E(f_\rho(x)-y)^2 )\\
& \leq & 5 C_K^2 M^2_\rho,
\end{eqnarray*}
\begin{eqnarray*}
& & \<f^\ast_\lambda, L_K (f^\ast_\lambda-f_\rho)\>_K \\
& \leq & \|L_K^{1/2} f^\ast_\lambda\|_K \cdot \| L_K^{1/2} (f^\ast_\lambda-f_\rho)\|_K \\
& = & \|f^\ast_\lambda\|_{\L^2_\rho} \cdot \|f^\ast_\lambda-f_\rho\|_{\L^2_\rho} \\
& \leq & 2 M_\rho^2 ,
\end{eqnarray*}
and $\lambda^2 \|f^\ast_\lambda\|^2_K \leq \lambda M^2_\rho$.
Thus we end the proof.
\end{pf}

@@

In this paper, we have shown by probabilistic upper bounds that a two-stage online learning algorithm,
the stochastic approximation of the gradient descent method followed by an averaging process, can achieve
the almost sure convergence with an optimal asymptotic rate with respect to the sample size, as good as ``batch learning''.
Moreover considering the regularization parameter and confidence, the best results obtained so far are,
$O(\lambda^{-2} t^{-1/2} \log^{1/2} 1/\delta)$ (Theorem B) or $O(\lambda^{-1} t^{-1/2} \delta^{-1/2})$ (Theorem B*).

@@

2. For $\alpha\in (0,1]$, $\theta\in [0,1)$, and all $t\in \N$,
\[ \psi^0_\theta(t,k,\alpha):=\sum_{j=k}^{t-1} \prod_{i=k}^j \left(1 - \frac{\alpha}{i^\theta} \right)
\leq \frac{D_\theta-1}{\alpha} k^{\theta};
\]

@@

2. Notice that
Let $\alpha'=\alpha/(1-\theta)$. Using part 1 with $p=1$, we obtain
\begin{eqnarray*}
\prod_{i=k}^j \left(1 - \frac{\alpha}{i^\theta} \right) & \leq &  e^{\alpha'[k^{1-\theta} - (j+1)^{1-\theta}]},
\end{eqnarray*}
whence
\begin{eqnarray*}
& & \sum_{j=k}^{t-1} \prod_{i=k}^j \left(1 - \frac{\alpha}{i^\theta} \right) \\
& \leq & \sum_{j=k}^{t-1} e^{\alpha'[k^{1-\theta} - (j+1)^{1-\theta}]} \\
& \leq & e^{\alpha' k^{1-\theta}}\int_{k}^\infty e^{- \alpha' x^{1-\theta}} d x \\
& \leq & (\alpha')^{-1/(1-\theta)} C_\theta' [(\alpha')^{1/(1-\theta)} k]^{\theta}, \\
& & \ \ \ \ \ \mbox{(by Lemma \ref{lem:key2})} \\
& \leq & 2^{\theta/(1-\theta)} \left\{ 1+\Gamma\left(\frac{1}{1-\theta}\right)\right\} \left(\frac{1}{\alpha}\right)k^\theta \\
& = & \frac{D_\theta-1}{\alpha} k^\theta.
\end{eqnarray*}
